---
title: "Analysis"
author: "Henry Oliver"
date: November 30, 2025
format: html
---

### Purpose
The purpose of this analysis is to provide maps and statistics that summarize the total area of suitable aquaculture habitat for various species of farmed aquatic species. First, the extent of aquaculture habitat for oysters in 5 west-coast economic zones will be determined. Then, I employ a function which allows a user to input ranges for depth in meters and temperature in degrees Celsius for a target species, and yield a similar map and table which summarizes approximate viable aquaculture area for each economic zone.

### Prepare environment - load libraries
```{r}
#| message: false
library(tidyverse)
library(sf)
library(terra)
library(tmap)
library(patchwork)
library(kableExtra)
```


### Read in Data
```{r}
#| message: false
# Create path for sst data
sst_dir <- here::here("data", "sst")
# Read in data as a stack of rasters
sst <- list.files(sst_dir, pattern = glob2rx("*.tif$"), full.names = TRUE)
sst_sort <- sst[
  # Sort filepaths based on numeric suffix
  order(
    # Extract numeric suffix of filenames and convert to numeric
    as.numeric(gsub(".*_(\\d+)\\.tif$", "\\1", sst))
    )
  ]
sst_rast <- rast(sst_sort)

# Read in bathymetry raster
bathymetry <- rast(here::here("data", "depth.tif"))

# Read in EEZ shapefile
eez <- st_read(here::here("data", "wc_regions_clean.shp"), quiet = TRUE)
```


### Calculate mean surface temperature in degrees Celsius
```{r}
# Calculate the mean sea surface temperature, adjust from K to C
mean_sst <- mean(sst_rast) - 273.15

# Reproject sst to a CRS with meters instead of degrees, specify resolution
sst_reprojected <- project(mean_sst, "EPSG:32611")

# Check values
sst_reprojected
```


### Adjust bathymetry data and combine with reprojected SST data
```{r}

bathymetry_reprojected <-  project(bathymetry, sst_reprojected) %>% # Reproject
  resample(sst_reprojected) %>%  # Resample
  crop(sst_reprojected) # Crop to sst extent

```

### Create matrices for relcassification

We want to isolate the temperatures and ocean depths that have been identified as suitable aquaculture habitats for our targeted species. In order to do this, we must re-classify our temperature and depth data to add an indicator (1 or 0) or whether a position on the raster is viable (1) or non-viable (0) for aquaculture of the targeted species. Once we've created re-classified sst and depth rasters, we'll combine them into a single suitable-habitat raster.
```{r}
# Create a matrix to reclassify SST
sst_matrix <- matrix(c(
            -Inf, 11, 0,
            11, 30, 1,
            30, Inf, 0),
ncol = 3,
byrow = TRUE
)
# Matrix for depth
depth_matrix <- matrix(c(
            -Inf, -70, 0,
            -70, 0, 1,
            0, Inf, 0),
ncol = 3,
byrow = TRUE
)
```

### Reclassify
```{r}
# Reclassify
bath_mask <- classify(bathymetry_reprojected, depth_matrix)
sst_mask <- classify(sst_reprojected, sst_matrix)

# Multiply, return raster where both bathymetry and sst masks have a value of 1
oyster_mask <- bath_mask*sst_mask
```

### Prepare EEZ data
```{r}
# Reproject
eez_proj <- st_transform(eez, crs(oyster_mask))

# Convert to SpatVector
eez_vect <- vect(eez_proj)

# Clip oyster habitat to eez extent with mask
oyster_eez <- mask(oyster_mask, eez_proj)
```

### Sum overlapping EEZ and habitat pixel area, print result
```{r}
# Get cellsizes in m^2
oyster_cell_sizes <- cellSize(oyster_eez, unit = "m")

# Get area from cell sizes
oyster_area <- oyster_eez*oyster_cell_sizes

# Get cellsizes of entire raster in km^2
total_area_km2 <- global(oyster_area, fun = "sum", na.rm = TRUE)[1,1] / 1e6

print(paste("Oyster suitable habitat is", round(total_area_km2), "km^2"))

```
### Create map showing suitable area to check our area summary function

```{r}
#| message: false
habitat_map <-
tm_shape(oyster_eez) +
  tm_raster(
    col.scale = tm_scale_categorical(
      values = c("0" = "lightgray", "1" = "darkgreen"),
      labels = c("Not Suitable", "Suitable")
    ),
    col.legend = tm_legend(title = "Oyster Habitat")
  ) +
  tm_scalebar(position = c("left", "bottom")) +
  tm_compass(position = c(0.9, 0.20)) +
  tm_layout(
    main.title = "Suitable Habitat for Oyster Farming",
    frame = FALSE
  )

habitat_map
```


### Create map and table of suitable habitat per eez zone
```{r}
#| message: False

# Suitable habitat per zone in km^2
zone_area <- zonal(oyster_area, eez_vect, fun = "sum", na.rm = TRUE) / 1e6

# Add column to eez_proj with the suitable areas from zone_area
eez_suitable_area <- eez_proj %>% 
  mutate(suitable_area = zone_area$depth)


eez_zones_map <-
  tm_shape(eez_suitable_area) +
  tm_polygons(
    fill = 'suitable_area',
    fill.scale = tm_scale_continuous(values = "viridis"),
    fill.legend = tm_legend(title = "Suitable Area (km^2)")
  ) +
  tm_compass(position = c(0, 0.2)) +
  tm_scalebar(position = c(0, 0.075)) +
  tm_title(text = "Suitable Habitat by Zone")

eez_zones_map

```

```{r results ='asis'}
# Create table
eez_zones_table <-
st_drop_geometry(eez_suitable_area) %>% 
  select(c('rgn', 'suitable_area')) %>% 
  kable(
    format = "html",
    col.names = c("Region", "Suitable Area (sq km)"),
    caption = "Suitable Area by EEZ Region for American cupped oyster",
    digits = 2
  )

eez_zones_table

```
### Interpretation of Figures

The map and table above show that Central and Southern California have the most suitable area for Oyster aquaculture, whereas Northern California and Oregon have notably lower suitable area for Oyster aquaculture.  This figure would suggest that aquaculture efforts for oysters would be best suited for the southern portion of the American West Coast.


### Function for Chosen Species - Pacific calico scallop **Argopecten ventricosus**
Here we will employ our 'aquaculture_range' function that takes temperature range and depth range for a target species, and produces similar figures to those generated above for Oysters. 

```{r}
#| message: false
# Source function
source(here::here("functions", "aquaculture_habitat_function.R"))

# Check Function for Pacific calico scallop
aquaculture_range(11, 28.5, -180, 0, species_name = "Argopecten ventricosus")

```

### Interpretation of Figure for *Argopecten ventricosus*

The figure and table generated show that there is significantly more suitable aquaculture habitat for Pacific calico scallops *Argopecten ventricosus* than there is for oysters on the West Coast of the US. This is likely because the scallops survivable depth range is more than double that of oysters. Furthermore, water off the coast of Oregon state contains nearly 10x more suitable habitat than the zone directly south, Northern California. 


### References

Hall, S. J., Delaporte, A., Phillips, M. J., Beveridge, M., & O’Keefe, M. (2011). 
*Blue Frontiers: Managing the Environmental Costs of Aquaculture*. 
The WorldFish Center, Penang, Malaysia.

Gentry, R. R., Froehlich, H. E., Grimm, D., Kareiva, P., Parke, M., Rust, M., 
Gaines, S. D., & Halpern, B. S. (2017). Mapping the global potential for marine aquaculture. 
*Nature Ecology & Evolution*, 1, 1317–1324.

SeaLifeBase. “United States (contiguous states) – Species summary for Argopecten ventricosus.” Accessed November 30, 2025. https://www.sealifebase.ca/country/CountrySpeciesSummary.php?Country=United+States+(contiguous+states)&genusname=Argopecten&speciesname=ventricosus

GEBCO Compilation Group (2022). *GEBCO_2022 Grid* 
(doi:10.5285/e0f0bb80-ab44-2739-e053-6c86abc0289c).


